---
title: "StatInf Course Project Part 1"
author: "Egor Ignatenkov"
date: "27.10.2014"
output: pdf_document
---
##Distribution of averages of exponentials


In this paper we explore the properties of the distribution of the mean of $n = 40$ exponentionals. The rate parameter $\lambda$ is $0.2$, and that means that mean of our exponentional distribution is $\mu=1/\lambda$, i.e. $5$ and the standard deviation is also $\sigma=1/\lambda$, i.e. $5$.

First, let's simulate our data. We will have $10000\times40$ matrix filled with exponentionally randomly generated values with rate parameter 0.2.

```{r}
n<-40
nosim<-10000
lambda<-0.2
sim<-matrix(rexp(nosim*n,lambda),nosim)
```

Now, let's create a vector with means of all the rows. It will be our sample from the distribution of the means.

```{r}
means<-apply(sim,1,mean)
```

Central Limit Theorem tells us that this distribution should be normal with mean $\mu = 1/\lambda = 5$ and standard deviation $\sigma/\sqrt n = 5/\sqrt 40 = 0.79$. Since we were asked about variance, square that: $\sigma^2/n = 25/40 = 0.625$. Let's see what we've really got.

###Mean of means

First, let's look on mean of our distribution:

```{r}
mean(means)
```

Seems pretty close to $5$.

###Variance

Now, variance:

```{r}
sd(means)^2
```

Not so far from $0.625$.

###Compare with normal distribution

Let's compare the looks of our sample and normal distribution. 

```{r}
library(ggplot2)
g<-qplot((means-mean(means))/sd(means),geom="density")
g<-g+stat_function(fun = dnorm, size = 2)
g
```

Again, quite close.

### Coverage of the confidence interval

We were also asked to evaluate the coverage of the confidence interval for $1/\lambda: \bar X \pm 1.96 S/\sqrt{n}$

```{r}
left<-mean(means)-1.96/(lambda*sqrt(n))
right<-mean(means)+1.96/(lambda*sqrt(n))
sum(means>left & means < right)/length(means)
```
And we almost hit 95% confidence interval, as we should have! Thank you very much.