---
title: "ToothGrowth"
author: "Egor Ignatenkov"
date: "27.10.2014"
output: pdf_document
---

Here we'll try to analyze data from ToothGrowth dataset.

First, let's load data and see how it looks.

```{r}
data(ToothGrowth)
str(ToothGrowth)
```

Let's factorize the dose column:
```{r}
ToothGrowth$dose<-as.factor(ToothGrowth$dose)
head(ToothGrowth)
summary(ToothGrowth)
```

##Let's see how growth depends on supp:

```{r}
g1<-ToothGrowth$len[ToothGrowth$supp == "VC"]
g2<-ToothGrowth$len[ToothGrowth$supp == "OJ"]
difference<-g2-g1
t.test(difference)
```

Confidence interval is all above zero, so we can easily suggest that with supp = OJ teeth grow faster.

##Let's check on dose:

```{r}
d1<-ToothGrowth$len[ToothGrowth$dose == 0.5]
d2<-ToothGrowth$len[ToothGrowth$dose == 1]
d3<-ToothGrowth$len[ToothGrowth$dose == 2]
diff21<-d2-d1
diff31<-d3-d1
diff32<-d3-d2
```

###1 against 0.5:

```{r}
t.test(diff21)
```

Again, confidence interval above zero, so when dose is 1, teeth grow better than with dose = 0.5

###2 against 0.5:

```{r}
t.test(diff31)
```

again, confidence interval above zero.

###2 against 1:

```{r}
t.test(diff32)
```

again, confidence interval above zero.


I should've thought how these two factors (dose and supp) may overlap and what I can do to deal with that but I have 7 minutes till deadline, sorry :(.